---
title: "Data Analysis 01 Template"
author: "Josh Davidsburg, Brittany N. Gaddy"
date: "3/2/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, FAs police struggle to solve homicides, Baltimore residents see an ‘open season for killing’, from the Washington Post [As police struggle to solve homicides, Baltimore residents see an ‘open season for killing](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(dplyr)
library(lubridate)
```

## Load and Cleaning Data

[In this block of text, describe the source of the data, and write a basic data dictionary for data you are working with]

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
homicide_data <- read.csv("data/homicide-data.csv")
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted. City police made an arrest in 41 percent of homicides in 2014; last year (2018), the rate was just 27 percent, a 14 percentage point drop.]
* **Analysis summary**: [I confirmed the posts findings. I brought in the data, filtered out Baltimore and both 2014 and 107, then I found the arrest rates for both years.]

```{r}
# Put code to reverse engineer sentence here

#Check to see how many days had 0 homicides
homicide_by_date <- homicide_data %>%
  filter(city == "Baltimore") %>%
  filter(reported_date >= 20150101) %>%
  group_by(reported_date) %>%
  summarise(homicides_per_day = n())
  
#find the arrest rate in 2014  
baltimore_homicide_2014 <- homicide_data %>%
  filter(city == "Baltimore") %>% 
  mutate(reported_date = ymd(reported_date)) %>%
  filter(reported_date >= as.Date("2014-01-01") & reported_date <= as.Date("2014-12-31"))%>%
  group_by(disposition) %>%
  summarise(total =n())

arrest_rate_2014 <- baltimore_homicide_2014$total[1] / sum(baltimore_homicide_2014$total)*100
  
arrest_rate_2014

#find the arrest rate in 2017  
baltimore_homicide_2017 <- homicide_data %>%
  filter(city == "Baltimore") %>% 
  mutate(reported_date = ymd(reported_date)) %>%
  filter(reported_date >= as.Date("2017-01-01") & reported_date <= as.Date("2017-12-31"))%>%
  group_by(disposition) %>%
  summarise(total =n())

arrest_rate_2017 <- baltimore_homicide_2017$total[1] / sum(baltimore_homicide_2017$total)*100
  
arrest_rate_2017



# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5
* **Sentence text**: [Community leaders and residents say that leaves hundreds of families who have been robbed of a loved one without a chance at seeing justice done. Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest.]
* **Analysis summary**: [I was able to confirm The Post's findings. There were 1,005 homicides in Baltimore between the beginning of 2015 and 2016. Only 252 of those homicides were "closed by arrest." So between the beginning of 2015 and 2016, one out of every four Baltimore homicides resulted in an arrest.]


```{r}


# Put code to reverse engineer sentence here
Homicides_number <- homicide_data %>%
  select(reported_date,city,disposition)%>%
  filter(city =="Baltimore", reported_date >= 20150101)%>%
  summarise(number_of_homicides=n())
  
Homicides_arrests <- homicide_data %>%
  select(reported_date,city,disposition)%>%
  filter(city =="Baltimore", reported_date >= 20150101, disposition=="Closed by arrest")%>%
  summarise(number_of_homicides_arrests=n())
one_out_of_four<- (Homicides_number/Homicides_arrests)
# Display results of code below this codeblock
Homicides_number
Homicides_arrests
one_out_of_four
```

-30-