---
title: "Data Analysis 01 Template"
author: "Josh Davidsburg"
date: "3/2/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, FAs police struggle to solve homicides, Baltimore residents see an ‘open season for killing’, from the Washington Post [As police struggle to solve homicides, Baltimore residents see an ‘open season for killing](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(dplyr)
library(lubridate)
```

## Load and Cleaning Data

[In this block of text, describe the source of the data, and write a basic data dictionary for data you are working with]

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
homicide_data <- read.csv("data/homicide-data.csv")
# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted. City police made an arrest in 41 percent of homicides in 2014; last year (2018), the rate was just 27 percent, a 14 percentage point drop.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

#Check to see how many days had 0 homicides
homicide_by_date <- homicide_data %>%
  filter(city == "Baltimore") %>%
  filter(reported_date >= 20150101) %>%
  group_by(reported_date) %>%
  summarise(homicides_per_day = n())
  
#find the arrest rate in 2014  
baltimore_homicide_2014 <- homicide_data %>%
  filter(city == "Baltimore") %>% 
  mutate(reported_date = ymd(reported_date)) %>%
  filter(reported_date >= as.Date("2014-01-01") & reported_date <= as.Date("2014-12-31"))%>%
  group_by(disposition) %>%
  summarise(total =n())

arrest_rate_2014 <- baltimore_homicide_2014$total[1] / sum(baltimore_homicide_2014$total)*100
  
arrest_rate_2014

#find the arrest rate in 2017  
baltimore_homicide_2017 <- homicide_data %>%
  filter(city == "Baltimore") %>% 
  mutate(reported_date = ymd(reported_date)) %>%
  filter(reported_date >= as.Date("2017-01-01") & reported_date <= as.Date("2017-12-31"))%>%
  group_by(disposition) %>%
  summarise(total =n())

arrest_rate_2017 <- baltimore_homicide_2017$total[1] / sum(baltimore_homicide_2017$total)*100
  
arrest_rate_2017



# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-