---
title: "Data Analysis 01 Template"
author: "Julia Arbutus, Josh Davidsburg, Brittany Gady, Ashkan Motamedi"
date: "3/2/2021"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, As police struggle to solve homicides, Baltimore residents see an ‘open season for killing’, from the Washington Post [As police struggle to solve homicides, Baltimore residents see an ‘open season for killing’](https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#load the tidyverse
library(tidyverse)
library(dplyr)
library(lubridate)
library(janitor)
library(scales)
```

## Load and Cleaning Data

Our data set comes from the Washington Post's collection of more than 52,000 criminal homicides over the past decade in 50 of the largest American cities. 

Each row is a homicide victim. 

* uid | character | unique id for each homicide victim
* reported_date | numeric | date homicide was reported
* victim_last | character | last name of victim
* victim_first | character | first name of victim
* victim_race | character | race of victim
* victim_age | character | age of victim
* victim_sex | character | sex of victim
* city | character | name of city
* state | character | two letter state code 
* lat | numeric | latitude of homicide
* lon | numeric | longitude of homicide
* disposition | character | status of homicide case


```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
homicide_data <- read.csv("data/homicide-data.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted. City police made an arrest in 41 percent of homicides in 2014; last year (2018), the rate was just 27 percent, a 14 percentage point drop.
* **Analysis summary**: We confirmed The Post's findings. We brought in the data, filtered for Baltimore and sorted the table so the column names would be the end result of the homicide case (closed by arrest, etc.). Then, we calculated the arrest rate for both 2014 and 2017 to get 41% and 27%. 

```{r}
#Check to see how many days in the last three years have had homicides.
arrest_rate <- homicide_data %>%
  mutate(reported_date = ymd(reported_date)) %>%
  filter(city == "Baltimore") %>%
  mutate(year = year(reported_date)) %>%
  group_by(year, disposition) %>%
  summarise(total=n()) %>%
  pivot_wider(names_from=disposition, values_from=total) %>%
  clean_names() %>% 
  mutate(total_homicides=closed_by_arrest+closed_without_arrest+open_no_arrest) %>%
  mutate(arrest_rate = (closed_by_arrest/total_homicides)*100) %>%
  filter(year %in% c("2014", "2017"))
arrest_rate
```

### Sentence 2

* **Sentence text**: Of 50 of the nation’s largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis. In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9.
* **Analysis summary**: We were able to confirm parts of The Post's findings. We found that Chicago's homicide arrest rate dropped 21 percentage points between 2014 and 2017, Boston's dropped 12 points and St. Louis's dropped 9. But, instead of finding that Baltimore is one of 34 cities with homicide arrest rate changes below zero, we found that Baltimore is one of 35 cities with homicide arrest rate changes below zero. (Below zero indicating police make arrests less often than in 2014). The 35th city is Tampa, Florida.

```{r}
# calculate number of arrests
arrest_rate <- homicide_data %>%
  mutate(reported_date = ymd(reported_date)) %>%
  mutate(year = year(reported_date)) %>%
  group_by(city, year, disposition) %>%
  summarise(total=n()) %>%
  pivot_wider(names_from=disposition, values_from=total) %>%
  clean_names() %>% 
  mutate(closed_by_arrest = replace_na(closed_by_arrest,0),
         closed_without_arrest = replace_na(closed_without_arrest,0),
         open_no_arrest = replace_na(open_no_arrest,0)
         ) %>%
  mutate(total_homicides=closed_by_arrest+closed_without_arrest+open_no_arrest) %>%
  mutate(arrest_rate = (closed_by_arrest/total_homicides)*100) %>%
  select(city, year, arrest_rate) %>%
  filter(year %in% c("2014", "2017")) %>%
  pivot_wider(names_from=year,values_from=arrest_rate) %>%
  clean_names() %>%
  mutate(change_2014_2017 = x2017-x2014) %>%
  arrange(change_2014_2017)
arrest_rate
```

### Sentence 3

* **Sentence text**: Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population. While homicide rates remain near historical lows in most American cities, Baltimore and Chicago are now both seeing murder tallies that rival the early 2000s.
* **Analysis summary**: We were able to confirm part of The Post's findings. We found that 2014 was the year Baltimore led the nation in the greatest raw number increase in killings with Chicago, but we were not able to confirm the number of 30 cities that had seen an increase in homicides. We found 25 states that saw an increase from 2014 to 2017, which was the closest number to the 30 cities reported in the article.

```{r}
arrest_rate <- homicide_data %>%
  mutate(reported_date = ymd(reported_date)) %>%
  mutate(year = year(reported_date)) %>%
  group_by(city, year) %>%
  summarise(total=n()) %>%
  pivot_wider(names_from=year, values_from=total) %>%
  clean_names() %>% 
  mutate_if(is.numeric, replace_na, 0) %>%
  mutate(change_2014_2017 = x2017-x2014) %>%
  arrange(change_2014_2017)
arrest_rate
```

### Sentence 4

* **Sentence text**: While there is evidence for and against a nationwide Ferguson effect — the theory that crime increased after 2014 as police faced more scrutiny following the shooting of Michael Brown in Ferguson, Mo. — in Baltimore there is an indisputable Freddie Gray effect. As violence in the city has risen since 2015, the likelihood of a killer being arrested has dropped precipitously.
* **Analysis summary**: We confirmed The Post's findings. We found that from 2015, total homicides in Baltimore have increased, while the arrest rate has decreased. 

```{r}
arrest_rate_drop <- homicide_data %>%
  filter(city == "Baltimore") %>%
  mutate(reported_date = ymd(reported_date)) %>%
  mutate(year = year(reported_date)) %>%
  group_by(year, disposition) %>%
  summarize(total = n()) %>%
  pivot_wider(names_from=disposition, values_from=total) %>%
  clean_names() %>% 
  mutate(total_homicides = closed_by_arrest+closed_without_arrest+open_no_arrest) %>%
  mutate(arrest_rate = (closed_by_arrest/total_homicides)*100) %>%
  select(year, total_homicides, arrest_rate)
  
ggplot(arrest_rate_drop, aes(x = year, y = total_homicides)) + geom_line() + scale_x_continuous(breaks = pretty_breaks())
ggplot(arrest_rate_drop, aes(x = year, y = arrest_rate)) + geom_line() + scale_x_continuous(breaks = pretty_breaks())
arrest_rate_drop
```

### Sentence 5

* **Sentence text**: Community leaders and residents say that leaves hundreds of families who have been robbed of a loved one without a chance at seeing justice done. Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest.
* **Analysis summary**: We confirmed The Post's findings. We found that there were 1,002 homicides between 2015 and the beginning of 2017, and that only 252, or one out of four, were closed by arrest.

```{r}
Homicides_number <- homicide_data %>%
  select(reported_date,city,disposition)%>%
  filter(city =="Baltimore", reported_date >= 20150101)%>%
  group_by(disposition) %>%
  summarise(total=n()) %>%
  pivot_wider(names_from=disposition, values_from=total) %>%
  clean_names() %>%
  mutate(total_homicides=closed_by_arrest+closed_without_arrest+open_no_arrest) %>%
  mutate(arrest_rate = total_homicides/closed_by_arrest)
Homicides_number
```

-30-